---
title: "House Price Prediction - King County, WA"
author: "Steve Su, Pierson Wodarz"
date: "July 17, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description of data file

This data set contains variables which can used to predict the prices of houses in King County, WA.  Examples of predictors include, number of bathrooms/bedrooms, square footage of living space/basement, and number of floors which are all numberic data types.  This data set also contains several categorical variables which include, view/no view, zipcode, and waterfront property.  All together there are roughly 20 predictors with 21500 observations.  The response is house price.  The data source was from Kaggle, https://www.kaggle.com/swathiachath/kc-housesales-data. 

Price prediction is important in many fields but can be subjective espcially in real estate.  Having a model that can predict prices objectly would be benefical between sellers and buyers....?    


```{r}
df_hse = read.csv('kc_house_data.csv')
head(df_hse)
```

